---
title: "Palliative Care"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(stringr)
library(tidyr)
library(lmtest)
library(fastDummies)
library(misty)
library(tidyverse)
library(readxl)
```

Find out whether socioeconomic status, age, sex, and ethnicity have an impact on patient rates of palliative care consultation, infection, intubation (outcome variable), (days to vent), discharge, etc.
```{r, warning=FALSE}
demo <- read_csv("/stor/work/FRI-BigDataBio/seton_data_analysis/SetonData_Feb/demo.csv")
demo <- demo %>% select(-1)

consult_notes <- read_csv("/stor/work/FRI-BigDataBio/seton_data_analysis/SetonData_Feb/consult_notes.csv")
consult_notes <- consult_notes %>% select(-1)

diagnosis <- read_csv("/stor/work/FRI-BigDataBio/seton_data_analysis/SetonData_Feb/diagnoises.csv")
diagnosis <- diagnosis %>% select(-1)

diagnosis_code <- read_excel("/stor/work/FRI-BigDataBio/seton_data_analysis/diagnosis_code.xlsx")
diagnosis_code <- diagnosis_code %>% select(1:2)

oxygen <- read_csv("/stor/work/FRI-BigDataBio/seton_data_analysis/SetonData_Feb/oxygen.csv")
```

```{r}
demo_data <- demo %>% select(patient_id_deident, encounter_id_deident, facility_name, race, ethnicity, gender, age, financial_class, discharge_disp, vent_days, vent_max_dt_indexed, vent_min_dt_indexed)

consult_notes <- consult_notes %>% select(patient_id_deident, 
                                          encounter_id_deident, document_title)

palliative_consult <- consult_notes %>% 
  filter(grepl(paste('palliative'), tolower(document_title)))

palliative_consult <- distinct(palliative_consult, patient_id_deident,
                               .keep_all = TRUE)

palliative_consult <- palliative_consult %>% select(-patient_id_deident)

diagnosis_code_infection <- diagnosis_code %>% 
  filter(grepl(paste('infection'), tolower(`SHORT DESCRIPTION (VALID ICD-10 FY2022)`)))

infection_code <- list(diagnosis_code_infection$CODE)[[1]]

diagnosis <- diagnosis %>% mutate_at('diagnosis_code', str_remove_all, '\\.')

infection_data <- diagnosis %>% filter(diagnosis_code %in% infection_code) %>% select(encounter_id_deident, diagnosis_code, diagnosis_description)
```

```{r}
all_data <- demo_data %>% left_join(palliative_consult, by = 'encounter_id_deident')
all_data <- all_data %>% mutate('palliative' = case_when(is.na(document_title) ~ '0',
                                                    TRUE ~ '1'))

all_data <- all_data %>% left_join(infection_data, by = 'encounter_id_deident')
all_data <- all_data %>% mutate('infection' = case_when(is.na(diagnosis_code) ~ '0',
                                                    TRUE ~ '1'))

all_data <- all_data %>% mutate('intubation' = case_when(is.na(vent_days) ~ '0',
                                                    TRUE ~ '1'))

all_data <- all_data %>% mutate(outcome = NA)
for (i in 1:nrow(all_data)){
  if (grepl('discharged', tolower(all_data$discharge_disp[i]))){
    all_data$outcome[i] = 'discharged'
  }
  else if (grepl('expired', tolower(all_data$discharge_disp[i]))){
    all_data$outcome[i] = 'expired'
  }
}
```

race vs. palliative care
```{r}
# chi-square test
chi_data <- all_data %>% select(race, palliative)

chi_data <- chi_data %>% group_by(race) %>% dplyr::count(palliative) %>% 
  pivot_wider(names_from = palliative, values_from = n)

chi_data <- column_to_rownames(chi_data, 'race')

chi_data

chisq <- chisq.test(chi_data)

chisq

mosaicplot(chi_data, col = hcl(c(120, 10)), main = 'Race & Palliative Care')

# logistic regression
palliative_race <- all_data %>% select(patient_id_deident, race, palliative) %>% unique()
palliative_race <- palliative_race %>% filter(race != 'MISSING')
palliative_race <- dummy_cols(palliative_race)
palliative_race <- palliative_race %>% select(-c(race, palliative_0, palliative_1))
palliative_race$palliative <- as.factor(palliative_race$palliative)

palliative_glm <- glm(palliative ~ `race_BLACK OR AFRICAN AMERICAN` + race_OTHER, data = palliative_race, family=binomial()) 

summary(palliative_glm)

phi_result <- misty::cor.phi(palliative_race %>% select(-patient_id_deident))$result

phi_result <- as.data.frame(phi_result)

phi_result
```

ethnicity vs. palliative care
```{r}
# chi-square test
chi_data <- all_data %>% select(ethnicity, palliative)

chi_data <- chi_data %>% group_by(ethnicity) %>% dplyr::count(palliative) %>% 
  pivot_wider(names_from = palliative, values_from = n)

chi_data <- chi_data %>% filter(!is.na(ethnicity))

chi_data <- column_to_rownames(chi_data, 'ethnicity')

chi_data

chisq <- chisq.test(chi_data)

chisq

mosaicplot(chi_data, col = hcl(c(120, 10)), main = 'Ethnicity & Palliative Care')

# logistic regression
palliative_ethnicity <- all_data %>% select(patient_id_deident, ethnicity, palliative) %>% unique()
palliative_ethnicity <- palliative_ethnicity %>% filter((ethnicity != 'UNKNOWN') & (!is.na(ethnicity)))
palliative_ethnicity <- dummy_cols(palliative_ethnicity)
palliative_ethnicity <- palliative_ethnicity %>% select(-c(ethnicity, palliative_0, palliative_1))
palliative_ethnicity$palliative <- as.numeric(palliative_ethnicity$palliative)

palliative_glm <- glm(palliative ~ `ethnicity_HISPANIC OR LATINO`, data = palliative_ethnicity, family=binomial())

summary(palliative_glm)

phi_result <- misty::cor.phi(palliative_ethnicity %>% select(-patient_id_deident))$result

phi_result <- as.data.frame(phi_result)

phi_result
```

age vs. palliative care
```{r}
# logistic regression
palliative_age <- all_data %>% select(patient_id_deident, age, palliative) %>% unique()
palliative_age <- palliative_age %>% filter(!is.na(age))
palliative_age <- palliative_age %>% select(-patient_id_deident)
palliative_age$palliative <- as.numeric(palliative_age$palliative)

palliative_glm <- glm(palliative ~ age, data = palliative_age, family=binomial())

summary(palliative_glm)

# point biserial correlation
cor.test(palliative_age$age, palliative_age$palliative)
```

gender vs. palliative care
```{r}
# chi-square test
chi_data <- all_data %>% select(gender, palliative)

chi_data <- chi_data %>% group_by(gender) %>% dplyr::count(palliative) %>% 
  pivot_wider(names_from = palliative, values_from = n)

chi_data <- chi_data %>% filter(!is.na(gender))

chi_data <- column_to_rownames(chi_data, 'gender')

chi_data

chisq <- chisq.test(chi_data)

chisq

mosaicplot(chi_data, col = hcl(c(120, 10)), main = 'Race & Palliative Care')

# logistic regression
palliative_gender <- all_data %>% select(patient_id_deident, gender, palliative) %>% unique()
palliative_gender <- palliative_gender %>% filter(!is.na(gender))
palliative_gender <- dummy_cols(palliative_gender)
palliative_gender <- palliative_gender %>% select(-c(gender, palliative_0, palliative_1))
palliative_gender$palliative <- as.factor(palliative_gender$palliative)

palliative_glm <- glm(palliative ~ gender_F, data = palliative_gender, family=binomial()) 

summary(palliative_glm)

phi_result <- misty::cor.phi(palliative_gender %>% select(-patient_id_deident))$result

phi_result <- as.data.frame(phi_result)

phi_result
```

intubation vs. palliative care
```{r}
# chi-square test
chi_data <- all_data %>% select(intubation, palliative)

chi_data <- chi_data %>% group_by(intubation) %>% dplyr::count(palliative) %>% 
  pivot_wider(names_from = palliative, values_from = n)

chi_data <- column_to_rownames(chi_data, 'intubation')

chi_data

chisq <- chisq.test(chi_data)

chisq

mosaicplot(chi_data, col = hcl(c(120, 10)), main = 'intubation & Palliative Care')
```

Expired patients (stop index only)
```{r}
expired <- demo %>% filter(discharge_disp == 'EXPIRED') %>% 
  select(patient_id_deident, encounter_id_deident, disch_date_indexed, 
         death_date_indexed)

expired <- expired %>% mutate(death_date_indexed = case_when(is.na(death_date_indexed) ~ disch_date_indexed, TRUE ~ death_date_indexed))

expired_meds <- expired %>% left_join((administered_meds %>% 
                         select(encounter_id_deident, med_generic_display,
                                stop_dt_tm_indexed)),
                      by = 'encounter_id_deident')

expired_meds <- expired_meds %>% 
  filter(death_date_indexed - stop_dt_tm_indexed <= 2)

distinct(expired_meds, patient_id_deident, med_generic_display, .keep_all = TRUE) %>%
  count(med_generic_display) %>% arrange(-n) %>% head(6)
```

(start and stop index)
```{r}
expired <- demo %>% filter(discharge_disp == 'EXPIRED') %>% 
  select(patient_id_deident, encounter_id_deident, death_date_indexed)

expired_meds <- expired %>% left_join((administered_meds %>% 
                         select(encounter_id_deident, med_generic_display,
                                start_dt_tm_indexed ,stop_dt_tm_indexed)),
                      by = 'encounter_id_deident')

expired_meds <- expired_meds %>% 
  filter((death_date_indexed - stop_dt_tm_indexed <= 2) & 
           (death_date_indexed - start_dt_tm_indexed <= 2))

distinct(expired_meds, patient_id_deident, med_generic_display, .keep_all = TRUE) %>% count(med_generic_display) %>% arrange(-n) %>% head(15)
```

facility vs. palliative care
```{r}
# chi-square test
chi_data <- all_data %>% select(facility_name, palliative)

chi_data <- chi_data %>% group_by(facility_name) %>% dplyr::count(palliative) %>% 
  pivot_wider(names_from = palliative, values_from = n)

chi_data <- column_to_rownames(chi_data, 'facility_name')

chi_data

chisq <- chisq.test(chi_data)

chisq

mosaicplot(chi_data, col = hcl(c(120, 10)), main = 'Facility & Palliative Care')

# logistic regression
palliative_facility <- all_data %>% select(patient_id_deident, facility_name, palliative) %>% unique()
palliative_facility <- dummy_cols(palliative_facility)
palliative_facility <- palliative_facility %>% select(-c(facility_name, palliative_0, palliative_1))
palliative_facility$palliative <- as.factor(palliative_facility$palliative)
palliative_facility <- palliative_facility %>% select(-patient_id_deident)
palliative_facility <- palliative_facility %>% select(-3)

palliative_glm <- glm(palliative ~ ., data = palliative_facility, family=binomial()) 

summary(palliative_glm)

phi_result <- misty::cor.phi(palliative_facility)$result

phi_result <- as.data.frame(phi_result)

phi_result
```
